Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2021-09-20 analysis/paper_draft2.Rmd
updated

---
title: "Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data"
date: "`r Sys.Date()`"
author:
  - Sachin S. Pandya
#    orcid_id: 0000-0001-7387-1307
  - Xiaomeng Li
  - Timothy E. Moore
#    orcid_id: 0000-0002-9576-0517
#output: distill::distill_article
output: pdf_document
bibliography: references.bib
abstract: |
  US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, gender, or membership in certain other cognizable classes. Here, we propose a Bayesian approach for detecting such illegal bias. We show how to use historical data on an attorney's use of peremptory strikes in past cases to estimate, in a current trial, how likely an attorney has struck prospective jurors because of cognizable-class membership. We also use of a _power prior_ to ajust the weight of such historical data in the analysis. We show the utility of our model on simulated data. Finally, we extend our approach by presenting a prototype software application to be used to detect bias in peremptory-challenge in real time during jury-selection. We demonstrate the app using actual historical strike data from a convenience sample of cases from the federal district court of Connecticut from YYYY to YYYY.

@@

Since the US Supreme Court's decision in _Batson v. Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or membership in another cognizable class. The party bringing a _Batson_ challenge bears the burden of proving such illegal bias [@LaFave2020, sec. 22.3]. For similar challenges under State law, a few States accept proof not of motive but only that an "objective observer" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [e.g., Wash. R. 37(e)]. Peremptory-challenge procedure, and thus the task of proving illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].

In any case, evidence supporting such challenges may include aggregate data on peremptory challenges in past cases. For example, in a State criminal case, evidence for a _Batson_ challenge can include the "relevant history of the State's peremptory strikes in past cases." [@flowers2019, 2243]. Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strike data from a single case [e.g. @Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because jury-selection records are often not readily available [@Grosso2017; @Wright2018] and otherwise require painstaking effort to collect [e.g., @APM2018].

When collecting strike data in multiple trials, prior studies have aggregated strikes across trials, reported the observed difference in strike rates by race or gender, and tested for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference. [@Eisenberg2017, 334-339; see also @Gastwirth2014; @Gastwirth2013]. Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001]. For different approaches to the same historical strike data from from Mississippi's Fifth Circuit Court District (1992-2017), see @Craft2018a (logistic regression); @DeCamp2020 (propensity score matching); and @Dunn2021 (optimal feature selection, optimal classification trees).

In this paper, we take a Bayesian approach to _Batson_ challenges [@Kadane2018] and extend it to incorporate historical strike data. In this approach, we specify a model of the peremptory-strike process in the court of interest that includes bias parameter $b$, to which we assign an initial prior distribution. Then, we use an attorney's strike data from past trials, as generated by the same peremptory-strike process, to estimate a posterior distribution for $b$. In so doing, we use the _power prior_ [@Ibrahim2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.  Finally, we use that posterior distribution as the prior for $b$ in the present case and update the posterior based on data on the target attorney's strikes in that present case.

@@

Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)

$$
Pr(\delta_{ij} | b) = 
\begin{cases}
  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=1\\    
  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  
\end{cases}
$$

@@

$$Pr(\delta_{ij}| b) = (\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}$$

@@

$$ 
L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}
$$
where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.

@@

$$\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b)$$
where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.

In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2. This departs from @Kadane2018, who assigned $b$ a prior of $Unif(-6,6)$. We assume that the law assigning the burden of proof for _Batson_ challenges entails a weakly-informative prior. For our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge as well as the initial burden of producing some relevant evidence of illegal bias (the _Batson_ prima facie case). Thus, if that challenging party in fact produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven. This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data.

@@

$$
L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8})
$$

@@

$$L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}$$

is the likelihood function of $b$ given historical data. $n_{0i}$ is the total number of jury selections in the historical data, and $n_{0j}$ is the total number of peremptory strikes in the historical data. $\delta_{0ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class" in the historical trials.

@@

To do this, we generated strike data for a current trial using seven different values of the bias parameter ("current" $b = {-3,-2,-1,0,1,2,3}$. As introduced above, $b>0$ ($b<0$) represents a bias against (preference) for a prospective juror within (outside) the cognizable class, while $b=0$ denotes neither bias or preference. A greater absolute value of $b$ represents greater bias against or preference for the cognizable class.

We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ( _historical_ $b = {-3,-2,-1,0,1,2,3}$).

We let historical $b$ differ from current $b$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random. Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature that the current trial has but that all the past trials in the historical data do not. To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not. If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias $b$ in the current case with a Black defendant.

@@

The following figures depict the coverage rates of $b$ under the 80\% credible interval (Figure \@ref(fig:b80)), under the 90\% credible interval (Figure \@ref(fig:b90)), and under the 95\% credible interval (Figure \@ref(fig:b95)).

@@

p80 <- ggplot()+
  geom_tile(data = sims, aes(x = d, y = d_h,
                        fill = as.factor(sig.n)),
            color="black")+
  facet_grid(h~a, labeller = label_both)+
  # scale_fill_gradient2(expression(hat(d)), 
  #                      low = "darkred", mid = "white", 
  #                      high = "blue", 
  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
  #                      limits = c(-3.2, 3.2),
  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
  xlab("Current b")+
  ylab("Historical b")+
  scale_fill_manual(values = c("white", "blue"))+
  coord_equal() +
  theme_minimal() +
  labs(fill = "coverage") +
  theme(axis.text = element_text(size =10), 
        #axis.title = element_text(size = 16), 
        #legend.title.align = 0.05, 
        #legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14))
p80

@@

For any given case $i$ in which jury selection occurs and a party uses $s$ number of strikes under this process, @Barrett2007 models the probability of a party striking at least $x$ cognizable-class members at random as following a hypergeometric distribution:

$$Pr(X \geq x) = \sum_{x}^{s} \frac{{c_{i} \choose x} {m_{i} \choose s - x}}{{c_{i} + m_{i} \choose s}}$$

where $x = 0, 1, 2, 3, . . . s$; $c_{i}$ denotes the number of cognizable class members in the venire subject to strike; and $m_{i}$ denotes the number of cognizable class non-members in the venire subject to strike.

@@

load_path_b1 <- here::here("b1.csv")
load_path_b2 <- here::here("b2.csv")
load_path_b3 <- here::here("b3.csv")

@@

@@

b1.t$b = "1"
b2.t$b = "2"
b3.t$b = "3"
b.com <- rbind(b1.t, b2.t, b3.t)
b.com$sig <- ifelse(b.com$`2.5%`>0, "Bias", 
                                  "No Bias")
b.com$b.num <- as.numeric(b.com$b)

b1.mean <-data.frame(b_s = mean(b1.t$b_s), 
                     `2.5%` =  mean(b1.t$`2.5%`),
                     `97.5%` = mean(b1.t$`97.5%`),
                     sim = "mean", 
                     b = "1", 
                     sig = "No Bias", 
                     b.num = 1)
b2.mean <-data.frame(b_s = mean(b2.t$b_s), 
                     `2.5%` =  mean(b2.t$`2.5%`),
                     `97.5%` = mean(b2.t$`97.5%`),
                     sim = "mean", 
                     b = "2", 
                     sig = "Bias", 
                     b.num = 2)
b3.mean <-data.frame(b_s = mean(b3.t$b_s), 
                     `2.5%` =  mean(b3.t$`2.5%`),
                     `97.5%` = mean(b3.t$`97.5%`),
                     sim = "mean", 
                     b = "3", 
                     sig = "Bias", 
                     b.num = 3)

@@

simplot <- ggplot(data = b.com, aes(x = sim.n, y = b_s, color = sig))+
  facet_grid(~b)+
  geom_point(size = 3)+
  geom_errorbar(aes(x = sim.n, ymin = `2.5%`, 
                    ymax = `97.5%`), width = 0.4)+
  geom_hline(data = b.com, aes(yintercept = b.num), col = "red", lwd = 1.1) +
  geom_hline(yintercept = 0, col = "black", lwd = 1, lty = 1) +
  geom_vline(xintercept = 51.25, col = "black", lwd = 1, lty = 2) +
  xlab("Simulation")+
  scale_color_manual("Bias", values = c("blue", "grey"))+
  ylab("Posterior Bias Value")+
  ylim(c(-0.5, 5))+
  scale_x_continuous(breaks = c(1:50, 52.5), labels = c(1:50, "Average"))+
  coord_flip()+
  theme_classic()+
  theme(axis.title = element_text(size = 16), 
        axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 9))

@@

To illustrate, suppose the strike tally in Table \@ref(tab:inpts) represents the pattern of strikes in the present case with prosecutor "X" and defense attorney "Y". To detect racial bias in the use of peremptory strikes using both the strike tally and historical strike data, we set the cognizable-class to "Race"; select the names of the prosecutor and defense attorney from their respective pull-down menus; and set the weight setting to "Equal". The app will in turn display the posterior density plots in Figure \@ref(fig:brian).

```{r brian, fig.cap="Posterior Density Plot, a = 1, cognizable class = race"}
library(dplyr)
library(Rcpp)
load_path2 <- here::here("app/mh_sampler_pp.cpp")
load_path3 <- here::here("app/jury_data_cleaned_new.rds")
sourceCpp(load_path2)
sourceCpp(load_path2)
dat0 <- readRDS(load_path3)
df <- data.frame(round = c(1:10),
                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
                 total = c(15,14,13,12,11,10,9,8,7,6),
                 cog = c(0,1,1,0,0,1,1,0,0,0),
                 party = rep(c("PP","PD"),5))
extract_atny <- function(atny_name,pp,dat){
  if(pp){
    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
    return(dat[which(idx),])
  }
  else{
    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
    return(dat[which(idx),])
  }
} 
organize_input <- function(dat,pp,cog){
  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
  if(cog=='gender'){
    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
  }else{
    dat_strikes <- dat_strikes %>% filter(!is.na(race))
  }

@@

Now suppose that the strike tally indicates the pattern of strikes by these attorneys against women. If so, we can select "gender" as the cognizable class, and the density plots becomes Figure \@ref(fig:brian2).

```{r brian2, fig.cap="Posterior Density Plots, a = 1, cognitive class = gender"}

@@

Finally, if we give low weight on historical information, the density plot will be as shown in figure \@ref(fig:brian3).

```{r brian3, echo=FALSE,fig.cap="Posterior Density Plot, a = 0.2"}
load_path2 <- here::here("app/mh_sampler_pp.cpp")
load_path3 <- here::here("app/jury_data_cleaned_new.rds")
sourceCpp(load_path2)
sourceCpp(load_path2)
dat0 <- readRDS(load_path3)
df <- data.frame(round = c(1:10),
                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
                 total = c(15,14,13,12,11,10,9,8,7,6),
                 cog = c(0,1,1,0,0,1,1,0,0,0),
                 party = rep(c("PP","PD"),5))
extract_atny <- function(atny_name,pp,dat){
  if(pp){
    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
    return(dat[which(idx),])
  }
  else{
    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
    return(dat[which(idx),])
  }
} 
organize_input <- function(dat,pp,cog){
  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
  if(cog=='gender'){
    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
  }else{
    dat_strikes <- dat_strikes %>% filter(!is.na(race))
  }

@@

>>>>

---
title: "Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data"
date: "`r Sys.Date()`"
author:
  - Sachin S. Pandya
#    orcid_id: 0000-0001-7387-1307
  - Xiaomeng Li
  - Timothy E. Moore
#    orcid_id: 0000-0002-9576-0517
#output: distill::distill_article
output: 
  pdf_document:
    latex_engine: xelatex
    number_sections: true
    fig_width: 5
    fig_height: 5
    fig_caption: true
header-includes: 
  - \usepackage{amsmath}
fontsize: 12pt
bibliography: references.bib
abstract: |
  US law bars using peremptory strikes during jury selection because of prospective juror race, ethnicity, gender, or membership in certain other cognizable classes. Here, we propose a Bayesian approach for detecting such illegal bias. We show how to use historical data on an attorney's use of peremptory strikes in past cases to estimate, in a current trial, how likely an attorney has struck prospective jurors because of cognizable-class membership. We also use the _power prior_ to adjust the weight of such historical data in the analysis. We show the utility of our model on simulated data. Finally, we extend this with a prototype software application to detect bias in peremptory-challenge in real time during jury-selection. We demonstrate the app using actual historical strike data from a convenience sample of cases from the federal district court of Connecticut.

@@

Since _Batson v. Kentucky_ [-@batson1986], a party violates the Equal Protection Clause of the US Constitution by using peremptory challenges if motivated by the prospective jurors' race, ethnicity, gender, or membership in another cognizable class. The party bringing a _Batson_ challenge bears the burden of proving such illegal bias [@LaFave2020, sec. 22.3]. For similar challenges under State law, a few States require only that an "objective observer" or an "objectively reasonable person" would find that race, ethnicity or another cognizable-class membership was a "factor" in that party's use of strikes [e.g. @WashR37, subsec. (e); Cal. Code Civ. Proc. $\S$ 231.7(d)(1)]. Peremptory-challenge procedure, and thus the task of proving illegal bias, varies not only by State, but also by court, including the number of peremptory challenges assigned to each side and the order in which each party uses those strikes [@nscs-jury2018; @Williams2017].

In any case, evidence of such illegal bias may include data on a party's use of peremptory challenges in past cases. [@flowers2019, 2243; @WashR37, subsec. (g)(v); Cal. Code Civ. Proc. $\S$ 231.7(d)(3)(G)]. Prior studies have collected historical strike data in past cases and reported the observed difference in strike rates by race or gender. Typically, they test for the probability of observing a non-zero difference in strike rates by race or gender, given repeated sampling from a hypothetical population of peremptory strikes with zero such difference [e.g., @Eisenberg2017, 334-339; @Grosso2012, 1548; for discussion, see @Gastwirth2014; @Gastwirth2013]. Prior studies have also modeled how much a prospective juror's race affected the odds of being struck [e.g., @Baldus2001]. For different approaches using the same historical strike data from Mississippi, see @Craft2018a (logistic regression); @DeCamp2020 (propensity score matching); and @Dunn2021 (optimal feature selection, optimal classification trees).

In this paper, we extend a Bayesian approach to _Batson_ challenges [@Kadane2018] to incorporate historical strike data. In this approach, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter $b$, to which we assign an initial prior distribution. Then, we use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for $b$. In so doing, we use the _power prior_ [@Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter $b$.  Finally, we use that posterior distribution as the prior for $b$ in the present case and update the posterior based on data on the target attorney's strikes in that present case.

@@

Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation `\@ref(eq:model1)`

\begin{equation}
Pr(\delta_{ij} | b) = 
\begin{cases}
  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=1\\    
  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  
\end{cases} (\#eq:model1)
\end{equation}

@@

\[
Pr(\delta_{ij}| b) = (\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}
\]

@@

\begin{equation}  
L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}(\#eq:model2)
\end{equation}

where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes.

@@

\begin{equation}  
\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b)
\end{equation}

where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed.

In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2. This departs from @Kadane2018, who assigned $b$ a prior of $\text{Uniform}(-6,6)$. We assume that the law assigning the burden of proof for _Batson_ challenges entails a weakly-informative prior. For our initial prior of $N(0,2)$, we let $\mu = 0$, because the law assigns the burden of proof in a _Batson_ challenge to the party bringing the challenge. Thus, if the challenging party produces no relevant evidence of illegal bias, the law requires a trial judge to reject the challenge as unproven. This is tantamount to treating zero as the most-likely value of the bias parameter, absent any data.

@@

\begin{equation}
L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8}) (\#eq:model4)
\end{equation}

@@

\begin{equation}
L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}
\end{equation}

is the likelihood function of $b$ given historical data; $n_{0i}$ is the total number of jury selections (trials) in the historical data; $n_{0j}$ is the total number of peremptory strikes in the historical data; and $\delta_{0ij}$ denotes whether or not a party used that strike on a person who belongs to a cognizable class in the historical trials.

@@

To do this, we generated strike data for a current trial using seven different values of the bias parameter ("current" $b = {-3,-2,-1,0,1,2,3}$. As introduced above, $b>0$ ($b<0$) represents bias against (for) a prospective juror within (outside) the cognizable class, while $b=0$ denotes no bias.

We also generated historical strike data of three sizes (same, double, and triple the size of current data, i.e., data on one previous trial, two previous trials) using seven different values of the bias parameter for generating that historical data ( _historical b_$ $= {-3,-2,-1,0,1,2,3}$).

We let historical $b$ differ from current $b$ (our parameter of interest) to simulate noise associated with historical strike data that is incomplete not at random. Alternatively, this difference simulates a real difference in the bias parameter that depends on a feature that the current trial has but that all the past trials in the historical data do not. To illustrate, suppose prosecutor with a bias value conditional on defendant race: $b > 0$ if the criminal defendant is Black, $b = 0$ if not. If our historical strike data contains only in cases with non-Black defendants, accounting for that historical data may lead us to underestimate prosecutor's bias ($b$) in the current case with a Black defendant.

@@

Figures 1, 2, and 3 depict the coverage rates of $b$ under an 80\%, 90\%, 95\% credible interval, respectively.

@@

p80 <- ggplot()+
  geom_tile(data = sims, aes(x = d, y = d_h,
                        fill = as.factor(sig.n)),
            color="black")+
  facet_grid(h~a, labeller = label_both)+
  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
  xlab("Current b")+
  ylab("Historical b")+
  scale_fill_manual(values = c("white", "blue"))+
  coord_equal() +
  theme_minimal() +
  labs(fill = "coverage") +
  theme(axis.text = element_text(size =10), 
        #axis.title = element_text(size = 16), 
        #legend.title.align = 0.05, 
        #legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14))
p80

@@

@@

load_path_b1 <- here::here("data-sim", "b1.csv")
load_path_b2 <- here::here("data-sim", "b2.csv")
load_path_b3 <- here::here("data-sim", "b3.csv")

@@

library(dplyr)

@@

b1.t$b = "1"
b2.t$b = "2"
b3.t$b = "3"
b.com <- rbind(b1.t, b2.t, b3.t)
b.com$sig <- ifelse(b.com$`2.5%`> 0, "Bias", 
                                  "No Bias")
b.com$b.num <- as.numeric(b.com$b)

b1.mean <-data.frame(b_s = mean(b1.t$b_s), 
                     `2.5%` =  mean(b1.t$`2.5%`),
                     `97.5%` = mean(b1.t$`97.5%`),
                     sim = "mean", 
                     b = "1", 
                     sig = "No Bias", 
                     b.num = 1)

b2.mean <-data.frame(b_s = mean(b2.t$b_s), 
                     `2.5%` =  mean(b2.t$`2.5%`),
                     `97.5%` = mean(b2.t$`97.5%`),
                     sim = "mean", 
                     b = "2", 
                     sig = "Bias", 
                     b.num = 2)

b3.mean <-data.frame(b_s = mean(b3.t$b_s), 
                     `2.5%` =  mean(b3.t$`2.5%`),
                     `97.5%` = mean(b3.t$`97.5%`),
                     sim = "mean", 
                     b = "3", 
                     sig = "Bias", 
                     b.num = 3)

@@

simplot <- ggplot(data = b.com, 
                  aes(x = sim.n, y = b_s, color = sig))+
  facet_grid(~b)+
  geom_point(size = 2)+
  geom_errorbar(aes(x = sim.n, ymin = `2.5%`, 
                    ymax = `97.5%`), width = 0.4)+
  geom_hline(data = b.com, aes(yintercept = b.num), col = "red", lwd = 1.1) +
  geom_hline(yintercept = 0, col = "black", lwd = 1, lty = 1) +
  geom_vline(xintercept = 51.25, col = "black", lwd = 1, lty = 2) +
  xlab("Simulation")+
  scale_color_manual("Bias", values = c("blue", "grey"))+
  ylab("Posterior Bias Value")+
  ylim(c(-0.5, 5))+
  scale_x_continuous(breaks = c(1:50, 52.5), labels = c(1:50, "Average"))+
  coord_flip()+
  theme_classic()+
  theme(axis.title = element_text(size = 16), 
        axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 9))

@@

To illustrate, suppose the strike tally in Table \@ref(tab:inpts) represents the pattern of strikes in the present case with prosecutor "X" and defense attorney "Y". To detect racial bias in the use of peremptory strikes using both the strike tally and historical strike data, we set the cognizable-class to "Race"; select the names of the prosecutor and defense attorney from their respective pull-down menus; and set the weight setting to "Equal". The app will in turn display the posterior density plots in Figure `\@ref(fig:brian)`.

```{r brian, eval = FALSE, fig.cap="Posterior Density Plot, a = 1, cognizable class = race"}
library(dplyr)
library(Rcpp)
load_path2 <- here::here("app/mh_sampler_pp.cpp")
load_path3 <- here::here("app/jury_data_cleaned_new.rds")
sourceCpp(load_path2)
sourceCpp(load_path2)
dat0 <- readRDS(load_path3)
df <- data.frame(round = c(1:10),
                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
                 total = c(15,14,13,12,11,10,9,8,7,6),
                 cog = c(0,1,1,0,0,1,1,0,0,0),
                 party = rep(c("PP","PD"),5))
extract_atny <- function(atny_name,pp,dat){
  if(pp){
    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
    return(dat[which(idx),])
  }
  else{
    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
    return(dat[which(idx),])
  }
} 
organize_input <- function(dat,pp,cog){
  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
  if(cog=='gender'){
    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
  }else{
    dat_strikes <- dat_strikes %>% filter(!is.na(race))
  }

@@

Now suppose that the strike tally indicates the pattern of strikes by these attorneys against women. If so, we can select "gender" as the cognizable class, and the density plots becomes Figure `\@ref(fig:brian2)`.

```{r brian2, eval = FALSE, fig.cap="Posterior Density Plots, a = 1, cognitive class = gender"}

@@

Finally, if we give low weight on historical information, the density plot will be as shown in figure `\@ref(fig:brian3)`.

```{r brian3, eval = FALSE, fig.cap="Posterior Density Plot, a = 0.2"}
load_path2 <- here::here("app/mh_sampler_pp.cpp")
load_path3 <- here::here("app/jury_data_cleaned_new.rds")
sourceCpp(load_path2)
sourceCpp(load_path2)
dat0 <- readRDS(load_path3)
df <- data.frame(round = c(1:10),
                 num_cog = c(4,4,4,3,2,2,2,2,2,2),
                 total = c(15,14,13,12,11,10,9,8,7,6),
                 cog = c(0,1,1,0,0,1,1,0,0,0),
                 party = rep(c("PP","PD"),5))
extract_atny <- function(atny_name,pp,dat){
  if(pp){
    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$P_atty_l[x], fixed=TRUE))
    return(dat[which(idx),])
  }
  else{
    idx = sapply(1:nrow(dat),function(x) grepl(atny_name, dat$D_atty_l[x], fixed=TRUE))
    return(dat[which(idx),])
  }
} 
organize_input <- function(dat,pp,cog){
  dat_strikes <- dat %>% filter(!is.na(strike_seq)) 
  if(cog=='gender'){
    dat_strikes <- dat_strikes %>% filter(!is.na(sex))
  }else{
    dat_strikes <- dat_strikes %>% filter(!is.na(race))
  }

@@

Prior work on statistical methods for _Batson_ challenges, however, has largely focused on strike data from a single case [e.g. @Barrett2007; @Gastwirth2016; @Kadane2018; @Kadane2018a], perhaps because jury-selection records are often not readily available [@Grosso2017; @Wright2018] and otherwise require painstaking effort to collect [e.g., @APM2018].
